---
title: "400-figure-4"
author: "Mac Campbell"
date: "6/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning=FALSE, message=FALSE)
```

```{r}
library(tidyverse)
library(ggpubr)
library(gridExtra)
library(grid)
library(RcppCNPy)
library(ggrepel)
library(scatterpie)
library(viridis)
library(cowplot)
library(magick)
library(ggsn)
```

label size
```{r}
txt<-8
size<-18
```

### Generate PCA/Admix files
```{sh, eval=FALSE}
python $HOME/pcangsd/pcangsd.py -beagle /home/abenj93/projects/RelictDace/data/de_novo/RD_popgen/results_admix/GSJS.ST.SP_admix/RD_GSJS.ST.SP_glh.beagle.gz -kinship -o outputs/400/RD_GSJS.ST.SP.kinship -threads 10

python $HOME/pcangsd/pcangsd.py -beagle /home/abenj93/projects/RelictDace/data/de_novo/RD_popgen/results_admix/GSJS.ST.SP_admix/RD_GSJS.ST.SP_glh.beagle.gz \
-relate outputs/400/RD_GSJS.ST.SP.kinship.kinship.npy -admix -o outputs/400/RD_GSJS.ST.SP-kinremoved -threads 10
```


Keeping 257 individuals after filtering (removing 14)
Boolean vector of unrelated individuals saved as outputs/400/RD_GSJS.ST.SP-kinremoved.unrelated.npy (Binary)
Parsing Beagle file
Read 257 samples and 87711 sites

Estimating population allele frequencies
EM (MAF) converged at iteration: 14

Number of sites after MAF filtering (0.05): 85931

Estimating admixture with K=13, alpha=0, batch=10, seed=0


## PCA

```{r, echo=FALSE}
#' @param samples character vector with the individuals IDs in the order in which
#' they were passed in the bamlist to angsd.
#' @param cov covariance matrix
covar2pcs <- function(samples, cov) {
  
  
  eig <- eigen(cov, symm = TRUE)
  PC <- as.data.frame(eig$vectors) %>%
    as_tibble() %>%
    setNames(sprintf("PC-%02d", 1:ncol(.)))
  
  samtib <- tibble(sample = samples)
  
  list(
    PCs = bind_cols(samtib, PC),
    eigevalues = eig$values
  )
}
```

```{r}
data<-read_tsv("meta/location-codes.tsv")
bams<-read_tsv("bamlists/RD_GSJS.ST.SP.bamlist", col_names = c("Path"))
bams$Bam<-gsub("/home/abenj93/projects/RelictDace/data/de_novo/RD_alignment/RAD_align/","",bams$Path)
bams$Bam<-gsub("_R1.sort.proper.rmdup.bam","",bams$Bam)

bams<-bams %>% select(Bam) %>% rename(label=Bam) %>% mutate(Individual=1:n())

meta<-left_join(bams, data)

covk<-read_delim("outputs/400/RD_GSJS.ST.SP-kinremoved.cov", col_names=FALSE, delim=" ") %>% as.matrix()

#now we have 257 fish. We lost 14 due to relatedness filtering.... I don't know which ones that would be though.
#Boolean vector of unrelated individuals saved as RD_GSJS.ST.SP-kinremoved.unrelated.npy (Binary)
#I guess I'll have to go from that binary file.
excl<-npyLoad("outputs/400/RD_GSJS.ST.SP-kinremoved.unrelated.npy") %>% as_tibble()


#maybe the zeros are what we want?
#These we can subset or meta
kept<-meta$Individual[excl$value!=0]
dropped<-meta$Individual[excl$value==0]

#These are the ones that were dropped:
meta %>% filter(Individual %in% dropped)
```

```{r}
pcak <- covar2pcs(kept, covk)

pca_longk <- pcak$PCs %>%
  tidyr::gather(., key = "PC", "val", -sample)

# then expand a grid of the possible comparisons (ordered)
expgk <- expand.grid(sample = pcak$PCs$sample,
                    PCx = sprintf("PC-%02d", 1:6),
                    PCy = sprintf("PC-%02d", 1:6),
                    stringsAsFactors = FALSE) %>%
  tibble::as_tibble()

# then left join the pca results onto that
pca_pairsk <- dplyr::left_join(expgk, pca_longk, by = c("sample", "PCx" = "PC")) %>%
  dplyr::rename(val_x = val) %>%
  dplyr::left_join(pca_longk, by = c("sample", "PCy" = "PC")) %>%
  dplyr::rename(val_y = val)

npc <- 3
pck <- pca_pairsk %>%
  filter( (PCx %in% sprintf("PC-%02d", 1:npc)) & 
            (PCy %in% sprintf("PC-%02d", 1:npc)) ) %>%
  left_join(., meta, by = c("sample" = "Individual")) %>%
  mutate(group = Locality) 

eigk <- eigen(covk, symm = TRUE)
vark <-eigk$values/sum(eigk$values)
cumvark <-cumsum(eigk$values)/sum(eigk$values)

head(vark)
head(cumvark)

pc12k<-pck %>% filter( (PCx =="PC-01") & (PCy =="PC-02") )
centers12k<-pc12k %>% group_by(Locality) %>% mutate(MeanX=mean(val_x), MeanY=mean(val_y)) %>% select(Locality, MeanX, MeanY) %>% unique()


k12<-ggplot(pc12k, aes(x=val_x, y=val_y, color=Locality))+
  geom_point(size=2, alpha=0.75)+
  theme_bw()+
  theme(panel.grid=element_blank())+
  xlab(paste("PC1", " ", round((100*vark[1]),2), "%", sep = ""))+
  ylab(paste("PC2", " ", round((100*vark[2]),2), "%", sep = ""))+
  theme(legend.position = "")+
  #ggtitle("Related Individuals Excluded PCs 1 & 2")+
  scale_x_reverse() 
  #theme(plot.title = element_text(hjust = 0.5))

k12 +  geom_text_repel(data=centers12k, aes(x=MeanX, y=MeanY, label=Locality), color="black", fontface='bold', size=2, max.overlaps = Inf)
fig5a<-k12
```


Let's add lables for SPKR, SPSH, STRT & STWA, STPR, STMR, GSJS

```{r}
GSJS<-centers12k %>% filter(str_detect(Locality, "GSJS")) %>% ungroup() %>% mutate(Area="Johnson Springs") %>% group_by(Area) %>% mutate(x=mean(MeanX), y=mean(MeanY)) %>% 
  select(Area, x, y) %>% unique()
STPR<-centers12k %>% filter(str_detect(Locality, "STPR"))  %>% ungroup() %>% mutate(Area="STPR BT2 & FT1") %>% group_by(Area) %>% mutate(x=mean(MeanX), y=mean(MeanY)) %>% 
  select(Area, x, y) %>% unique()
others<-centers12k %>% filter(Locality %in% c("SPKR", "SPSH", "STMR", "STRT", "STWA")) %>%
  rename(Area=Locality, x=MeanX, y=MeanY)

labels<-bind_rows(GSJS,STPR,others)
fig4a<-k12 +  geom_text_repel(data=labels, aes(x=x, y=y, label=Area), color="black", fontface='bold', size=3, max.overlaps = Inf) +
  ggtitle(label = "A") +
  theme(plot.title = element_text(face="bold", size=16)) 

fig4a
```
## Admixture

```{r}
k13<-npyLoad("outputs/400/RD_GSJS.ST.SP-kinremoved.admix.Q.npy") %>% as_tibble() %>% rename(Q1=V1, Q2=V2, Q3=V3, Q4=V4, Q5=V5, Q6=V6, Q7=V7, Q8=V8, Q9=V9, Q10=V10, Q11=V11, Q12=V12, Q13=V13) %>% mutate(Individual=1:n()) %>% mutate(Index=1:n())

keepers<-as_tibble(kept) %>% rename(Individual=value)
keepersMeta<-left_join(keepers, meta) %>% select(-Individual)
#keepersMeta$`Collection Locality`<-gsub(pattern = "GSJS_.*", replacement="GSJS", keepersMeta$Locality)
keepersMeta$`Collection Locality`<-keepersMeta$Locality
k13df<-bind_cols(keepersMeta, k13)

q13s<-k13df  %>% arrange(`Collection Locality`) %>% mutate(Index=1:n()) %>% dplyr::select(label, Locality, `Collection Locality`, Index, Q1, Q2, Q3, Q4, Q5, Q6, Q7, Q8, Q9,Q10,Q11,Q12,Q13) %>% gather(key=Ancestry, value=Q, 5:17)

q13s$`Collection Locality`<-as.factor(q13s$`Collection Locality`)

#need some levels
levels<-colnames(k13 %>% select(-Individual, -Index))
q13s$Ancestry<-factor(q13s$Ancestry, levels=levels)
```


```{r}
pops13<-q13s %>% group_by(`Collection Locality`) %>% mutate(Start=min(Index), Stop=max(Index)) %>% 
  select(`Collection Locality`,Start,Stop) %>% unique() %>% 
  mutate(Position=round((Start+Stop)/2)) %>% arrange(`Collection Locality`)
```


```{r}
pops2<-pops13
kplot13<-ggplot(q13s) +
  geom_col(aes(x=Index, y=Q, fill=Ancestry), color="NA", size = 0, width = 1)+
#  scale_fill_manual(values=c("blue","red")) +
  geom_segment(data=pops2, x = pops2$Start - 0.5, y=0, xend = pops2$Start-0.5, yend=1, alpha=0.9, size=0.25) +
  geom_segment(data=pops2, x = pops2$Stop[length(pops2$Stop)]  + 0.5, y=0, xend= pops2$Stop[length(pops2$Stop)] + 0.5, yend=1,  alpha=0.9,
               size=0.25) +
  geom_segment(x=0, xend= pops2$Stop[length(pops2$Stop)], y=1, yend=1, alpha=0.9, size=0.25) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  ylim(0,1.01) +
  xlim(-0.1, pops2$Stop[length(pops2$Stop)]+1) +
  theme(panel.background = element_blank())+
  scale_x_continuous(breaks=pops2$Position, labels=pops2$`Collection Locality`) +
  xlab("") +
  theme(axis.text.x = element_blank(), axis.ticks.x=element_blank()) +
  scale_x_continuous(breaks=pops2$Position, labels=pops2$`Collection Locality`) +
  xlab("Sampling Location") +
#  theme(legend.position="NA") +
  theme(axis.text.x = element_text(size=6, angle=45, hjust=1)) 

kplot13
```

## Plot of GSJS
Scatterpies don't use tidy data.

```{r}
total<-left_join(pc12k, k13df)
toplot<-total %>% select(-sample, -PCx, -PCy, -label, -Locality, -group, -Individual, -Index)
```

Let's look at GSJS node pies placed with coordinates and average pops.

```{r}
lats<-read_csv("meta/dace-and-snail-distribution.csv") %>% mutate(`Collection Locality`=paste0("GSJS_",Spring)) %>%
  select(-RelictDace, -SpringSnail, -Notes)
```

```{r}
newplot<-toplot %>% left_join(lats) %>% na.omit()
sum<-newplot %>% select(-Spring) %>% group_by(`Collection Locality`) %>% summarise_all(list(mean))
springs <- newplot %>% select(`Collection Locality`, Spring) %>% unique()
sum<-left_join(sum, springs)
sum$`Collection Locality`<-as.factor(sum$`Collection Locality`)
```

Get background imagery:

```{r}
image<-image_read("outputs/400/crop.png")
image2 <- image_fill(image, 'none')
raster <- as.raster(image2)
Longitude<-c(-114.7,-114.3)
Latitude<-c(40.73,41.2)
corners<-cbind(Longitude,Latitude) %>% as_tibble()
```

```{r}
pies<-ggplot(corners, aes(x=Longitude, y=Latitude)) +
  geom_point() +
  annotation_raster(raster, xmin = -114.525017-.001, xmax = -114.472267, ymin =  40.952688, ymax =  40.989213) +
  geom_scatterpie(data=sum %>% select(-val_x, -val_y, Spring), 
                    aes(x=Longitude, y=Latitude, group=`Collection Locality`, r=.0008), 
                           cols=colnames(toplot %>% select(-val_x, -val_y, -`Collection Locality`)),
                           color=NA) +
  geom_text_repel(data=sum, aes(x=Longitude, y=Latitude, label=Spring), color="white", fontface="bold", size=3) +
  theme_bw() +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "") +
  coord_fixed(ratio=1, xlim=c(-114.522, -114.495), ylim=c(40.96,40.987)) +
  ggsn::scalebar(x.min = -114.522, x.max= -114.495, y.min=40.96, y.max=40.987,
                 dist = .5, dist_unit = "km", st.size=3, height=0.02, transform=TRUE, model = 'WGS84',
                 st.color="black")


pies
```


Get colors the same:
https://stackoverflow.com/questions/8197559/emulate-ggplot2-default-color-palette
```{r}
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
```

```{r}
n=13
cols = gg_color_hue(n)
dev.new(width = 4, height = 4)
plot(1:n, pch = 16, cex = 2, col = cols)

```



```{r}
qs<-colnames(toplot %>% select(-val_x, -val_y, -`Collection Locality`))
q13st<-q13s
q13st$Ancestry<-factor(q13st$Ancestry, levels=qs)

k13plot<-ggplot(q13st) +
  geom_col(aes(x=Index, y=Q, fill=Ancestry), color="NA", size = 0, width = 1)+
#  scale_fill_manual(values=c("blue","red")) +
  geom_segment(data=pops2, x = pops2$Start - 0.5, y=0, xend = pops2$Start-0.5, yend=1, alpha=0.9, size=0.25) +
  geom_segment(data=pops2, x = pops2$Stop[length(pops2$Stop)]  + 0.5, y=0, xend= pops2$Stop[length(pops2$Stop)] + 0.5, yend=1,  alpha=0.9,
               size=0.25) +
  geom_segment(x=0, xend= pops2$Stop[length(pops2$Stop)], y=1, yend=1, alpha=0.9, size=0.25) +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  ylim(0,1.01) +
  xlim(-0.1, pops2$Stop[length(pops2$Stop)]+1) +
  theme(panel.background = element_blank())+
  xlab("") +
  theme(axis.text.x = element_blank(), axis.ticks.x=element_blank()) +
  scale_x_continuous(breaks=pops2$Position, labels=gsub("RD6","",pops2$`Collection Locality`)) +
  xlab("Sampling Location") +
#  theme(legend.position="NA") +
  theme(axis.text.x = element_text(size=6, angle=45, hjust=1)) 

k13plot
```

```{r}

fig4b<-k13plot+scale_fill_viridis(option="plasma", discrete=TRUE) + theme(legend.position = "") +
  theme(axis.title.x = element_blank()) +
  ggtitle(label = "B") +
  theme(plot.title = element_text(face="bold", size=16)) +
  theme(axis.text.x = element_text(size=txt))
fig4b
```

```{r}
fig4c<-pies+scale_fill_viridis(option="plasma", discrete=TRUE) + 
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  guides(fill=guide_legend(title="Ancestry Proportion")) +
  theme(legend.key.size = unit(.5, 'cm')) +
  theme(legend.text = element_text(size=6)) +
  theme(legend.title = element_text(size=8)) +
    ggtitle(label = "C") +
  theme(plot.title = element_text(face="bold", size=16)) 

fig4c
```

## Plot

```{r}
ggarrange(arrangeGrob(fig4a,fig4c, ncol=2, widths=c(1,1.5)), fig4b,
          ncol = 1, nrow = 2, widths = c(3,3), heights=c(20,12))

ggsave("outputs/400/Figure-4.jpg", width=8.5, height=11/2)
ggsave("outputs/400/Figure-4.pdf", width=8.5, height=11/2)

```

## Fst and Theta Pi
```{r}

RD_east = read.table("outputs/400/eastern_valley.fst.tsv", header = T, sep = '\t')
RD_east$Pop1<-gsub("STPR","STPR_",RD_east$Pop1)
RD_east$Pop2<-gsub("STPR","STPR_",RD_east$Pop2)

RD_east$Pop1<-factor(RD_east$Pop1, levels=c("GSTS","GSJS_BS","GSJS_CS02","GSJS_CS03","GSJS_CS04","GSJS_CS11","GSJS_CS16","GSJS_CS22A","GSJS_NS03","GSJS_NS05","GSJS_SS19","SPKR","SPSH","STMR","STPR_BT2","STPR_FT1","STRT"))

fst<-ggplot(RD_east, aes(Pop2, Pop1)) + 
  geom_tile(aes(fill = Fst), color="white") + 
  theme_minimal() + 
  scale_fill_viridis(discrete=FALSE, option = "cividis", space = "Lab", direction = -1) + 
  theme_bw() + 
  theme(axis.text.x=element_text(angle=45,vjust=1,size=txt,hjust=1), panel.border=element_blank(),
        panel.grid.major=element_blank(), axis.title.y = element_blank(), axis.title.x = element_blank()) + scale_y_discrete(position = "left") + 
  theme(axis.text.y=element_text(size=txt)) +
#  geom_text(aes(label = round(Fst, 2)), size=3.5, colour = "white") + 
  ggtitle("D") + 
  theme(plot.title = element_text(face="bold", size=16)) 
 
fst
```

```{r}
RD_east = read.table("outputs/400/eastern_valley.diversity.tsv", header = T, sep = '\t')
RD_east[,c(3,4)] = RD_east[,c(3,4)]/RD_east$nSites
RD_east = subset(RD_east, tW < 0.020 & tP < 0.020)
RD_east_means=aggregate(RD_east[,3:5], list(RD_east$Pop), FUN=mean, header=TRUE)


pi <- ggplot(RD_east, aes(x=Pop, y=tP)) + 
  stat_boxplot(geom ='errorbar') +
  geom_boxplot(aes(fill=Pop), notch=FALSE, outlier.size=0.15, outlier.stroke = 0, outlier.alpha=0.5) + 
  labs(y = "Theta Pi") + 
  labs(x="") +
  theme_bw() +
  scale_y_continuous(breaks=seq(0.00,0.05,0.0025)) + 
  ggtitle(label = "E") +
  theme(plot.title = element_text(face="bold", size=16)) +
  theme(panel.grid = element_blank()) +
  theme(legend.position = "") +
  #scale_fill_manual(values=c("#D55E00","#D55E00","#E69F00","#009E73","#0072B2","#0072B2")) +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1, size=txt), axis.title.x=element_blank()) +
  theme(axis.text.y = element_text(size=txt+4)) +
  theme(axis.title = element_text(size = size))
  #scale_x_discrete(labels=c("BTOC",BTQP","RBFL","RBFR","P226","P246"))

pi
```

## Plot
```{r}
blank <- grid.rect(gp=gpar(col="white"))
```

```{r}
g<-ggarrange(arrangeGrob(arrangeGrob(fig4a+theme(axis.title = element_text(size = size)) +
                                       theme(axis.text = element_text(size=txt+4)),blank, ncol=1, nrow=2, heights =c(9,1)),
                         fig4c+theme(axis.title = element_text(size = size)) +
                           theme(axis.text = element_text(size=txt+4)), ncol=2, widths=c(1,1.5)), 
             arrangeGrob(fig4b+theme(axis.title = element_text(size = size)) + 
                               theme(axis.text.y = element_text(size=txt+4)),blank, ncol=1, heights=c(8,2)),
          ncol = 1, nrow = 2, widths = c(3,3), heights=c(20,12))

g
```

```{r}
h<-ggarrange(fst, pi, ncol=1, nrow=2)
h
```

```{r}
ggarrange(g, h, ncol=2, widths=c(1.5,1))
ggsave("outputs/400/Figure-4-new.jpg", width=18, height=8.5)
ggsave("outputs/400/Figure-4-new.pdf", width=18, height=8.5)
```

Making Supplemental Figure S4

```{r}
fst +
  geom_text(aes(label = round(Fst, 2)), size=3.5, colour = "white") + 
  ggtitle("")

ggsave("outputs/400/Supplemental Figure S4.jpg")
```
